Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-10-15 manuscripts/supplement.Rmd
old commits

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    fig_crop: false
    template: components/elsarticle.latex
    keep_tex: true

@@

Appendix: Model definitions and estimation
===========================================

@@

>>>>

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    template: components/elsarticle.latex
    keep_tex: true

@@

Figure S1 illustrates the performance of the GP and parametric models
outside the observed training data. The mean trajectory under the
underlying model is shown by the black dots, while the corresponding
prediction made by the model shown by the box and whiskers plots.
Predictions are based on the true expected value in the previous time
step.  Predicted distributions that lie entirely above the expected
dynamics indicate the expectation of stock sizes higher than what is
actually expected. The models differ both in their expectations and
their uncertainty (colored bands show two standard deviations away).
Note that the GP is particularly uncertain about the dynamics relative
to structurally incorrect models like the Ricker.

```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP. Points show the stock size simulated by the true model.  Overlay shows the range of states predicted by each model, based on the state observed in the previous time step. The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).  Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
y <- numeric(8)
y[1] <- 4.5
for(t in 1:(length(y)-1))
      y[t+1] = z_g() * f(y[t], h=0, p=p)
# predicts means, does not reflect uncertainty estimate!
crash_data <- step_ahead_posteriors(y)
crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
ggplot(crash_data) + 
  geom_boxplot(aes(as.factor(as.integer(time)), value, 
                   fill = variable, col=variable), 
               outlier.size=1, position="identity") + 
#  geom_line(aes(time, value, col = variable, 
#            group=interaction(L1,variable))) + 
  geom_point(aes(time, stock), size = 3) + 
  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  facet_wrap(~variable) + 
  theme(legend.position="none") + xlab("time") + ylab("stock size")

write.csv(crash_data, "components/data/figure3.csv")
```

@@

Further sensitivity analysis
=============================

The Latin hypercube approach systematically varies all combinations of
parameters, providing a more general test than varying only one parameter
at a time.  We loop across eight replicates of three different randomly
generated parameter sets for each of two different generating models
(Allen and Myers) over two different noise levels (0.01 and 0.05),
for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process performs
nearly optimally in each case, relative to the optimal solution with no
parameter or model uncertainty (Figure S10, appendix).
